---
title: "superblock"
output:
  md_document:
    variant: markdown_github

  rmarkdown::html_vignette:
    self_contained: no
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r opts, echo = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE,
    warning = TRUE,
    message = TRUE,
    width = 120,
    comment = "#>",
    fig.retina = 2,
    fig.path = "man/figures/README-"
)
```

[![R-CMD-check](https://github.com/UrbanAnalyst/superblock/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/UrbanAnalyst/superblock/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/UrbanAnalyst/superblock/graph/badge.svg)](https://app.codecov.io/gh/UrbanAnalyst/superblock)


# superblock

An R package to analyse potential
[superblocks](https://doi.org/10.1016%2Fj.cities.2024.105609). The following
example demonstrates the main functionality.

## Example

An example of the "Hansaviertel" of [Münster,
Germany](https://www.openstreetmap.org/#map=17/51.955569/7.639795). Note that
this entire document is dynamically-generated from the four lines of code shown
immediately below. All numerical values described in the text are then taken
from the actual data as downloaded from Open Street Map (OSM) at the time the
document was generated.

The first steps are to load the library and extract the necessary data:

```{r extract-data-fakey, message = FALSE, eval = FALSE}
library (superblock)
bbox <- c (7.63099, 51.95048, 7.66402, 51.96142)
hw_names <- c ("Bremer Straße", "Bremer Platz", "Wolbecker Straße", "Hansaring")
osmdat <- sb_osmdata_extract (bbox, hw_names)
```

```{r extract-data-for-real, message = FALSE, echo = FALSE}
library (superblock)
bbox <- c (7.63099, 51.95048, 7.66402, 51.96142)
hw_names <- c ("Bremer Straße", "Bremer Platz", "Wolbecker Straße", "Hansaring")
outer <- c (T, F, T, T)
osmdat <- sb_osmdata_extract (bbox, hw_names, outer = outer)
```

The `sb_summary()` function then summarises change in publicly accessible area if
the area is converted to a superblock:

```{r sb-summary}
sb_summary (osmdat)
```


---

### Analysis of parking times

There are two additional functions to estimate how long people spend driving
around searching for car parking spaces, and how far they actually drive.

#### Individual times search for parking spaces

The first function estimates the time taken on average for each resident of the
superblock to find a car parking space, as a function of the proportion of
available car parking spaces. Results are based on simulations using actual
data for the specified superblock to estimate how long it takes on average to
fill all of the remaining parking spaces.

The results are compared to equivalent times to drive to the nearest large
parking facility ("garage" in the graph below), to park a car, and then walk
back to the place of residence.

```{r parking-times, echo = FALSE, warn = FALSE}
times <- sb_parking_times (osmdat, ntrials = 10, n_props = 20, prop_min = 0.7)
plot (times)
```


```{r parking-times-fakey, eval = FALSE}
times <- sb_parking_times (osmdat)
plot (times)
```

The results shows how long it takes to find a parking space for
an average attempt ("park(50%)"), for one out of every four
attempts ("park(75%)"), and for one out of every ten attempts
("park(90%)"). Above around 97-98% occupancy, which is the general state most
nights, an "average" attempt may still take under 5 minutes, yet one out of
four times will take 10 minutes or more, and one out of ten times it will take
more than half an hour. The horizontal lines shows the time it would take on
average for all residents of the area to drive directly to a nearby large
parking facility ("garage"), park, and walk back. Beyond around 95% occupancy,
it is always quicker to park in a garage and walk back.

#### Total distances driven while searching for parking spaces

The average times to find parking spaces shown above are only high in extreme
cases (the 90% line), and when almost all parks are full. Nevertheless, most
parking spaces are indeed completely full most nights of the week, suggesting
the some people must experience these extremes most days. An additional
function estimates the total distances driven by everybody while filling all
parking spaces.

```{r parking-per-km}
d <- sb_parking_km_per_empty (osmdat)
plot (d$prop_full, d$d,
    type = "l", col = "red", lwd = 2,
    xlab = "Initial proportion filled",
    ylab = "Distance per empty car park until completely filled (in km)"
)
```

```{r num-parking-spaces, echo = FALSE, warning = FALSE}
net <- superblock:::parking_as_dodgr_net (osmdat)
nspaces <- sum (net$np)
nspaces_filled <- round (0.3 * nspaces)

# Then get lengths of streets:
names <- unique (net$name [which (!is.na (net$name))])
net$flow <- 1
net1 <- dodgr::merge_directed_graph (net)
lens <- vapply (names, function (n) {
    index <- which (net1$name == n)
    sum (net1$d [index])
}, numeric (1L))
lens_within <- lens [which (!names (lens) %in% osmdat$hw_names)]
max_len <- sprintf ("%.1f", max (lens) / 1000)
max_len_within <- sprintf ("%.1f", max (lens_within) / 1000)
ten_drivers <- round (10 * (max (lens) + max (lens_within)) / 1000)
```


Because filling the final few spaces takes so long, and people have to drive so
far to fill them, the overall distance driven to fill all spaces is actually
quite long. The block has around `r nspaces` parking spaces in total, so with
an initial proportion filled of 70%, people have to drive a total of around 1km
times 0.3 &times; `r nspaces` = `r nspaces_filled`km. While this may seem
extreme, consider that most of this distance is covered by the last few people
driving to find parking spaces, who are forced to drive several kilometres.
The longest street within the block is `r max_len_within`km, while one of the
surrounding streets is `r max_len`km. Ten people driving along both of those
streets will cover `r ten_drivers`km, which is
`r round(100*ten_drivers/nspaces_filled)`% of the total distance of
`r nspaces_filled` km covered by that same number of drivers.

Or in terms of time, if the average driving speed while searching for a
parking space is 20km/hr, then those 186 people each take an average of
1/20th of an hour, or only 3 minutes, to find a park, yet aggregated
over all free parking spaces, that amounts to a total of
`r nspaces_filled / 20` hours for everybody to fill all remaining car
parks.
