---
title: "superblock"
output:
  md_document:
    variant: markdown_github

  rmarkdown::html_vignette:
    self_contained: no
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r opts, echo = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE,
    warning = TRUE,
    message = TRUE,
    width = 120,
    comment = "#>",
    fig.retina = 2,
    fig.path = "README-"
)
```

# superblock

An R package to analyse potential [superblocks](https://doi.org/10.1016%2Fj.cities.2024.105609).

## Example

An example of the "Hansaviertel" of [Münster,
Germany](https://www.openstreetmap.org/#map=17/51.955569/7.639795). First load
the library and extract the necessary data from Open Street Map (OSM):

```{r extract-data}
library (superblock)
bbox <- c (7.63099, 51.95048, 7.66402, 51.96142)
hw_names <- c ("Bremer Straße", "Bremer Platz", "Wolbecker Straße", "Hansaring")
outer <- c (T, F, T, T)
osmdat <- sb_osmdata_extract (bbox, hw_names, outer = outer)
```

These are OSM id values for a few roads which have no parking specified, yet
have cars parked.
```{r parking-ids}
add_parking_osm_ids <- c (
    "10876650", "10879386", "24420125", "43026134", "43026139", "50163329",
    "148585465", "283680476", "721322833", "721738382", "756777549", "1122509887"
)
```

Summarise change in publicly accessible area if converted to superblock:

```{r sb-summary}
sb_summary (osmdat, add_parking_osm_ids = add_parking_osm_ids)
```
